The finite base property for some relation algebras subreducts

### 2020-06-01 problem-19.17/Text.tex
update pdf
add bib files

fixes

update

update

fixes

\begin{defin}
  A residuated semigroup is an algebra $\mathcal{S} = \langle S, \cdot, \leq, \setminus, / \rangle$ such that $\langle S, \cdot, \leq, \rangle$ is an ordered residuated semigroup ${\bf ORS}$ an the following equivalences hold for each $a, b, c \in S$:

  \begin{center}
    $b \leq a \setminus c \Leftrightarrow a \cdot b \leq c \Leftrightarrow a \leq c / b$
  \end{center}
\end{defin}

It is clear that ${\bf ORS}$ is a quasivariety and, thus \cite{mal2012algebraic}, is closed under subalgebras, products, and ultraproducts, i. e.,
${\bf ORS} = {\bf S}{\bf P}{\bf Up} \: {\bf ORS}$.

@@

\begin{theorem}
  That might follow from the previous theorem, Theorem~\label{ors=irrs}, and something else.
\end{theorem}

@@

The hypothetical plan is the following one:
\begin{enumerate}
\item Define properly relativised representation for residuated semigroups, that should look like ternary Kripke frames for the basic Lambek calculus or arrow logic.
\item Define the standard translation to such first-order relation structures. TODO: take a look at the loosely guarded fragment stuff.
\item Every finite residuated semigroup has a finite relativised representation.
\item If every $\Pi_1$-statement $\varphi$ of the language of residuated semigroups that is valid in every
residuated semingroup is valid in algebra having a finite relativised representation (one may use here Theorem~\ref{ors=irrs} somehow), then $\varphi$ is valid in ${\bf ORS}$ as well as in ${\bf IRRS}$.
\item Every finite residuated semigroup should have a finite relativised representation.
\item Construct a finitely based relation residuated semigroup from that (an analogue of complex algebra or smth like that). This item is the most non-trivial one.
\item As a corollary, the first-order universal first-order theory of ${\bf IRRS}$ should be decidable and
(it seems so) $\operatorname{NP}$-complete (that should follow from the results in \cite{pentus2006lambek}).
\end{enumerate}

@@

\subsection{The finite base property for ${\bf RJSOS}$}

\subsection{The finite algebra property for ${\bf RJSOS}$}

\subsection{The finite algebra on finite base for ${\bf RJSOS}$}

>>>>

\begin{defin}
  A residuated semigroup is an algebra $\mathcal{S} = \langle S, \cdot, \leq, \setminus, / \rangle$ such that $\langle S, \cdot, \leq, \rangle$ is an ordered residuated semigroup and the following equivalences hold for each $a, b, c \in S$:

  \begin{center}
    $b \leq a \setminus c \Leftrightarrow a \cdot b \leq c \Leftrightarrow a \leq c / b$
  \end{center}
  ${\bf ORS}$ is the class of all residuated semigroups.
\end{defin}

@@

\begin{proof}
  That might follow from the previous theorem, Theorem~\ref{ors=irrs}, and something else.
\end{proof}

@@

The hypothetical plan is the following one:
\begin{enumerate}
\item Define properly relativised representation for residuated semigroups, that should look like ternary Kripke frames for the basic Lambek calculus or arrow logic.
\item Define the standard translation to such first-order relation structures. TODO: take a look at loosely guarded fragment stuff.
\item Every finite residuated semigroup has a finite relativised representation.
\item If every $\Pi_1$-statement $\varphi$ of the language of residuated semigroups that is valid in every
residuated semingroup is valid in algebra having a finite relativised representation (one may use here Theorem~\ref{ors=irrs} somehow), then $\varphi$ is valid in ${\bf ORS}$ as well as in ${\bf IRRS}$.
\item Every finite residuated semigroup should have a finite relativised representation.
\item Construct a finitely based relation residuated semigroup from that (an analogue of complex algebra or smth like that). This item is the most non-trivial one.
\item As a corollary, the first-order universal first-order theory of ${\bf IRRS}$ should be decidable and
(it seems so) $\operatorname{NP}$-complete (that should follow from the results in \cite{pentus2006lambek}). The Lambek calculus is decidable that was shown syntactically via cut elimination and subformula property.
Here we would have an alternative way of showing decidability for some substructural logics.
\end{enumerate}

@@

\subsection{The finite algebra on finite base for ${\bf RJSOS}$ (or its failure)}
